paper1_figures

### 2023-09-12 paper1_markdown.Rmd
12.09: first version of the "group identity" section

keywords          : "cooperation, social dilemmas, commons"
#wordcount         : ""

@@

classoption       : "man"
output            : papaja::apa6_word
---

@@

@@

@@

One of the most observed moderator of this effect is the interaction partners. In fact, in a social dilemma experiment, women are more cooperative toward strangers than men are, while men cooperate more in a group of friends. Moreover, after a socialization phase where strangers could interact, men cooperated more than women [@Peshkovskaya2017]. These results seem to indicate that men accord more importance at the relationship with other members than women. @Gomez-Ruiz2020 also got to the same conclusion by showing that the creation of a team identity enhanced men's cooperation, but had no effects on women. Moreover, in a social dilemma with many iterations, men, but not women, tended to increase their cooperation over time [@Balliet2011].

@@

@Gomez-Ruiz2020: even if study about gender differences, they found a main effect of team identity on cooperation
@Arora2016: baseline = reciprocity: participants respond with defection to defection and with cooperation to cooperation, but in-group identification can change this --> main effect of identification on cooperation
faced to an in-group defector, high identifiers cooperate more than low (no difference between high and low when the partner cooperates) = compensation + this compensation comes with lower trust toward the partner who defected, so the cooperation is not driven by in-group favoritism
@Torsvik2011: group identity as mediator of anticipating a discussion (see if useful or not to add it here)
@VanVugt2009: mention that it happens that high-identifying members compensate for free-riders + when people identify with a social group the are more concerned about their in-group reputation
boost of social identity through inducing competition between groups, but not helpful --> promote a superordinate social identity
group member that strongly identify with their community do not need an economic incentive to cooperate
@Puurtinen2009: between group competition enhances cooperation (even between unrelated individuals) by aligning individual and group interest --> investements in public good higher in the group competition vs normal condition + in the group competition there was more anger toward in-group free-rider and guilt toward its own lack of cooperation --> ingroup identification can be enhanced via intergroup competition
@Noonan2016: study about commons = dog parks --> "having a strong sens of community is related to increased voluntarism" + re-read to see if others parts could be interesting
@Zhang2019: collective risk as operationalization of group identity: (1) cooperation decreased with time when the risk was individual, but not when it was collective; (2) beliefs about others' cooperation formed similarly under the two conditions = these beliefs are interpreted in a different way according to the condition
@Dorrough2015: ingroup favoritism increase with time and it is mediated by expectations of others' contributions + the increment is due to the change between ingroup and outgroup as interaction partners, it is not gradual

@Bardsley2005: paper about conformity (see if it can be useful in this section)

#D

>>>>

keywords          : "cooperation, social dilemmas, commons"
wordcount         : ""

@@

classoption       : "man"
output            : papaja::apa6_word
#bibliography: references.bib
---

@@

@Torsvik2011: in the introduction says that people in social dilemma are confused and can make random choices = argument against the completely rational human (see wich papers they cited)

@@

-   @Kramer1984: see the phrases underlined in the introduction

@@

One of the most observed moderator of this effect is the interaction partners. In fact, in a social dilemma experiment, women were more cooperative toward strangers than men, while men cooperated more in a group of friends. Moreover, after a socialization phase where strangers could interact, men cooperated more than women [@Peshkovskaya2017]. These results seem to indicate that men accord more importance at the relationship with other members than women. @Gomez-Ruiz2020 also got to the same conclusion by showing that the creation of a team identity enhanced men's cooperation, but had no effects on women. Moreover, in a social dilemma with many iterations, men, but not women, tended to increase their cooperation over time [@Balliet2011].

@@

Social identity have a consensual positive effect on cooperation: the more individuals feel a strong identification with a group, the more they cooperate with other members of the group (e.g. @Arora2016; @Noonan2016; @Gomez-Ruiz2020). For example, in a field study about community-organized dog parks in the United States, @Noonan2016 found that a strong sense of community was related to a higher amount of voluntary activities to maintain the parks in good conditions. This is explained by the fact that, when people strongly identify with a social group, they are more concerned about their group's well-being and reputation [@VanVugt2009] and they accord more importance to group-related goal [@Arora2016]. A strong social identity also maintains a high level of cooperation over time [@Zhang2019]. Moreover, independently of the strength of group identification, people cooperate more with other people in the same group than with out-group people [@Dorrough2015].

However, the fact that not all members identify with their group to the same extent can lead to inequalities on cooperation. For example, @VanVugt2009 mentioned that high identified individual can even compensate, by cooperating more, for other in-group members' defection. The same result was also found by @Arora2016: high identified individuals cooperated more than low identified ones when faced to an in-group participant who defected (this difference was not found when the defector was not a group member). However, this compensation came with a cost, since participants who compensated for an in-group defector reported lower trust toward this person.

Another well replicated result is the positive effect of intergroup competition on intragroup cooperation. In fact, in-group identification can be enhanced via a competition between different groups because in such a context individual's and group's interests are aligned [@Puurtinen2009]. In a social dilemma experiment, these authors demonstrated that cooperation was higher in an intergroup-competition condition than in a non-competitive condition. Moreover, when participants had to compete against other groups, they reported higher levels of anger toward in-group free-riders and they felt more guilty after a defection decision. In another experiment, intergroup competition reduced resources consumption (and therefore enhanced cooperation), while individual, dispositional competitiveness predicted an higher level of harvesting [@Nockur2020]. This shows that the increase of cooperation via competition is due to group membership factors, and not to competition per se.

However, despite the observed beneficial effects of inducing an intergroup competition, in a real-life situation this could be harmful. This is especially true in case of resources that are shared between several groups, which are at greater risk of over-exploitation in case of intergroup competition [@VanVugt2009]. A possible solution to maintain the positive effects of group identification, while eliminating the risk of over-exploitation caused by intergroup competition, is to prime a superordinate identity [@VanVugt2009]. In fact, intergroup cooperation can be enhanced by making salient the common characteristics of the different groups, instead of their differences [@Kramer1984]. Another effective way to create a superordinate identity is to induce the feeling of a common fate, for example through a collective risk. In fact, a common fate allows a group to become real [@Zhang2019], and can therefore erase intergroup competition.

Finally, social identity does not only have a main effect on people's cooperation, but can also moderate other factors' impact. For example, monetary incentive are more effective at enhancing cooperation of low identified individual than of high identified ones [@VanVugt2009]. Moreover, a previous group failure can decrease cooperation between low identified member, but this effect is not observed in people who have a strong identification with their group [@Jackson2011; @Jackson2012].

In general, social identity seems to have a positive and largely observed effect on individuals' cooperation. This is due to the fact that humans have a very developed sense of belonging to social groups, and that this has an impact on their willingness to help other members of their group [@VanVugt]. Since priming a common social identity is quite simple (e.g. @Kramer1984; @Zhang2019), this could be a very powerful method to enhance cooperation in an easy and cheap way.

# Discussion
